---
title: 'Final Project'
date: "12/03/2021"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
options(width = 80)
```

# Load library

```{r}
library(dplyr)
library(ggplot2)
library(ggcorrplot)
library(GGally)
library(alr4)
library(pROC)
library(ResourceSelection)
library(caret)
library(caTools)
theme_set(theme_bw())
```

# Read and exam the Dataset

```{r}
bank_df <- read.csv("BankChurners.csv")
bank_complete <- bank_df %>% select(-starts_with("Naive_Bayes"))
head(bank_complete)
```
```{r}
typeof(bank_complete$Attrition_Flag)
```

# Descriptive analytics

## Dependent Variable

```{r}
summary(bank_complete$Avg_Utilization_Ratio)
bank_complete %>% ggplot(aes(x=Avg_Utilization_Ratio))+geom_histogram(bins=25,fill='blue',alpha=0.5)+ggtitle("Histogram of Avg_Utilization_Ratio")
bank_complete %>% ggplot(aes(y=Avg_Utilization_Ratio))+geom_boxplot(fill='blue',alpha=0.5)+ggtitle("Boxplot of Avg_Utilization_Ratio")
```

the distribution of avg utilization ratio is right-skewed. We applied the sqrt transformation on the variable

```{r}
bank_complete %>% ggplot(aes(x=sqrt(Avg_Utilization_Ratio)))+geom_histogram(bins=25,fill='blue',alpha=0.5)+ggtitle("Histogram of sqrt_Avg_Utilization_Ratio")
bank_complete %>% ggplot(aes(y=sqrt(Avg_Utilization_Ratio)))+geom_boxplot(fill='blue',alpha=0.5)+ggtitle("Boxplot of sqrt_Avg_Utilization_Ratio")
```

## Categorical variables

### Gender

```{r}
t.test(Avg_Utilization_Ratio~Gender,data=bank_complete)
bank_complete %>% ggplot(aes(y=Avg_Utilization_Ratio,x=Gender,fill=Gender))+geom_boxplot(alpha=0.7)+ggtitle("Boxplot of Avg_Utilization_Ratio by Gender")
```

### Education_Level
```{r}
education_anova <- aov(Avg_Utilization_Ratio~Education_Level,data=bank_complete)
summary(education_anova)
```

Not significant

### Marital_Status

```{r}
Marital_anova <- aov(Avg_Utilization_Ratio~Marital_Status,data=bank_complete)
summary(Marital_anova)
TukeyHSD(Marital_anova)
bank_complete <- bank_complete %>% mutate(Married=ifelse(Marital_Status=='Married','Yes','No'))
t.test(Avg_Utilization_Ratio~Married,data=bank_complete)
bank_complete %>% ggplot(aes(y=Avg_Utilization_Ratio,x=Married,fill=Married))+geom_boxplot(alpha=0.7)+ggtitle("Boxplot of Avg_Utilization_Ratio by Married or not")
```

### Income_Category

```{r}
Income_anova <- aov(Avg_Utilization_Ratio~Income_Category,data=bank_complete)
summary(Income_anova)
TukeyHSD(Income_anova)
bank_complete %>% ggplot(aes(y=Avg_Utilization_Ratio,x=Income_Category,fill=Income_Category))+geom_boxplot(alpha=0.7)+ggtitle("Boxplot of Avg_Utilization_Ratio by Income")
```

### Card_Category

```{r}
Card_anova <- aov(Avg_Utilization_Ratio~Card_Category,data=bank_complete)
summary(Card_anova)
TukeyHSD(Card_anova)
bank_complete <- bank_complete %>% mutate(BlueCard=ifelse(Card_Category=='Blue','Yes','No'))
t.test(Avg_Utilization_Ratio~BlueCard,data=bank_complete) #most significant
bank_complete %>% ggplot(aes(y=Avg_Utilization_Ratio,x=BlueCard,fill=BlueCard))+geom_boxplot(alpha=0.7)+ggtitle("Boxplot of Avg_Utilization_Ratio by Blue Card")
```

## Numerical Variables

```{r}
bank_num <- bank_complete%>% select(Customer_Age,Dependent_count, Months_on_book,Total_Relationship_Count,Months_Inactive_12_mon,
                        Contacts_Count_12_mon,Credit_Limit,Total_Revolving_Bal,Avg_Open_To_Buy,Total_Amt_Chng_Q4_Q1,
                        Total_Trans_Amt,Total_Trans_Ct,Total_Ct_Chng_Q4_Q1,Avg_Utilization_Ratio)
summary(bank_num)
```

### correlation between the average utilization ratio with all the potential numerical variables

```{r}
utilize_ratio_cor <- cor(select(bank_num,-Avg_Utilization_Ratio),bank_num$Avg_Utilization_Ratio,use='complete.obs')
utilize_ratio_cor
ggcorrplot(utilize_ratio_cor)+ggtitle("Correlation plot of avg_utilization_ratio")
```

Based on the correlation, we can find that Credit Limit, Total_Revolving_Bal, and Avg_Open_to_Buy have the highest correlation with dependent variable Avg_Utilization_Ratio

### Credit_Limit

```{r}
bank_num %>% ggplot(aes(x=Credit_Limit))+geom_histogram(bins=25,fill='blue',alpha=0.5)+ggtitle("Credit Limit histogram")
bank_num %>% ggplot(aes(x=Credit_Limit,y=Avg_Utilization_Ratio))+geom_point()+ggtitle("Avg Utilization ratio scatter plot by Credit Limit")
```

### Total_Revolving_Bal

```{r}
bank_num %>% ggplot(aes(x=Total_Revolving_Bal))+geom_histogram(bins=25,fill='red',alpha=0.6)+ggtitle("Total_Revolving_Balance histogram")
bank_num %>% ggplot(aes(x=Total_Revolving_Bal,y=Avg_Utilization_Ratio))+geom_point()+ggtitle("Avg Utilization ratio scatter plot by Total_Revolving_Balance")
```

### Avg_Open_To_Buy

```{r}
bank_num %>% ggplot(aes(x=Avg_Open_To_Buy))+geom_histogram(bins=25,fill='blue',alpha=0.4)+ggtitle("Avg_Open_To_Buy histogram")
bank_num %>% ggplot(aes(x=Avg_Open_To_Buy,y=Avg_Utilization_Ratio))+geom_point()+ggtitle("Avg Utilization ratio scatter plot by Avg_Open_To_Buy")
```

All three variables are moderately right-skewed, and we considered using transform on these variables.

# Multiple Linear Regression Analysis

## Full Model

```{r}
attach(bank_complete)
m_full<- lm(sqrt(Avg_Utilization_Ratio)~.-CLIENTNUM-Attrition_Flag-Education_Level-Marital_Status-Card_Category+log(Credit_Limit)+sqrt(Total_Revolving_Bal)+log(Avg_Open_To_Buy), data= bank_complete)
summary(m_full)
```

P-value of our full model is lower than 0.05. There is statistically sufficient evidence the model is valid. 

The adjusted R-squared is 0.9537, which means 95.37% of the variation in Y( sqrt(avg_utilization_ratio) can be explained by this model. 

However, Avg_open_to_Buy is showing NA in the model summary.

Firstly, Avg_open_to_Buy can be explained by the combination of the other variables including credit limit and Total revolving balance, which is the main reason it shows NA in the model result.

Secondly, since the utilization ratio is equal to total revolving balance divide by credit limit. So the data leakage would render our model meaningless.

## Model without total revolving balance and avg_open_to_buy

We removed both variables and refit the model

```{r}
m1 <- lm(sqrt(Avg_Utilization_Ratio)~.-CLIENTNUM-Attrition_Flag-Education_Level-Marital_Status-Card_Category+log(Credit_Limit)-Total_Revolving_Bal-Avg_Open_To_Buy, data= bank_complete)
summary(m1)
```

We find that the adjust r2 is dropping heavily, so we exam the diagnose plot

```{r}
par(mfrow=c(2,2))
plot(m1)
```

there is a line in both residuals vs Fitted plot and Scale-Location plot, and we determined that these are the data whose Avg utilization ratio is equal to 0

## Removed all the 0 value utilization ratio observation

```{r}
detach(bank_complete)

bank_complete_no0 <- bank_complete %>% filter(Avg_Utilization_Ratio>0)

attach(bank_complete_no0)
```

## Refit the model with 7657 observations with 16 variables

```{r}
m2<- lm(sqrt(Avg_Utilization_Ratio)~.-CLIENTNUM-Attrition_Flag-Education_Level-Marital_Status-Card_Category+log(Credit_Limit)-Total_Revolving_Bal-Avg_Open_To_Buy, data= bank_complete_no0)
summary(m2)
```


We applied the variable selection procedure to look for an optimal model. In order to make the model simple and efficient, we used the BIC criterion.

## Backward elimination model based on BIC

```{r}
n2 <- length(m2$residuals)
backBIC2 <- step(m2,direction = "back",data=bank_complete, k=log(n2))
summary(backBIC2)
```

## Stepwise selection model based on BIC

```{r}
stepBIC2 <- step(m2,direction = "both",data=bank_complete, k=log(n2))
summary(stepBIC2)
```

## Forward selection model based on the BIC criterion

```{r}
mint <- lm(sqrt(Avg_Utilization_Ratio)~1,data=bank_complete_no0)
forwardBIC <- step(mint,scope=list(lower=~1,upper=~Customer_Age+Dependent_count+Months_on_book+Total_Relationship_Count+
                                     Months_Inactive_12_mon+Contacts_Count_12_mon+Credit_Limit+ log(Credit_Limit)+
                                     Total_Amt_Chng_Q4_Q1+Total_Trans_Amt+ Total_Trans_Ct+Total_Ct_Chng_Q4_Q1+
                                     Gender+Married+Income_Category+BlueCard,direction = "forward", data=bank_complete_no0,k=log(n2)))
summary(forwardBIC)
```



Backward elimination and stepwise selection based on BIC choose the same model with 3 predictors, which are all significant in the model. And p-value is lower than 0.05, which means the model overall is highly significant. Because the Adjusted R-squared is 0.7852, it means 78.52% of the variation in the sqrt(avg_utilization_ratio) can be explained by the model

Only 5 variables are statistically significant in the forward selection model based on the BIC criterion, which has 6 variables.

Even though the forward selection model has 0.7858 adjusted R-squared, and can explain 0.06% extra variation in Y better than the stepwise model.

We would still omit this model, because it’s too complex.

## Regression diagnostics plot

```{r}
par(mfrow=c(2,2))
plot(stepBIC2)
```

We didn’t find any special pattern in residuals vs fitted plot, and we could see that 3943 is close to the cook’s distance bands, which means it is the bad leverage points; 5485,3943, and 4369 are the bad normality points because they are far away from the line in the Normal QQ plot..

We should investigate these points for further study.

# Logistic Regression

```{r}

bank_complete <- bank_complete%>%mutate(Attrition_Flag = factor(Attrition_Flag, levels = c('Existing Customer','Attrited Customer'),
                                                             labels = c(0,1)))
y <- bank_complete$Attrition_Flag
attach(bank_complete)
```

## Data visualization of dependent variable(Attrition_Flag) with all predictor variable

```{r}
par(mfrow=c(2,2))
boxplot(Customer_Age~y, ylab="Customer Age",xlab="Attrited Customer? (0=No, 1=Yes)")
boxplot(Dependent_count~y, ylab="Dependent Count",xlab="Attrited Customer? (0=No, 1=Yes)") #s
boxplot(Months_on_book~y, ylab="Months on Book",xlab="Attrited Customer? (0=No, 1=Yes)") #
boxplot(Total_Relationship_Count~y, ylab="Total Relationship Count",xlab="Attrited Customer? (0=No, 1=Yes)") #
boxplot(Months_Inactive_12_mon~y, ylab="Months Inactive 12 mon",xlab="Attrited Customer? (0=No, 1=Yes)")#s
boxplot(Contacts_Count_12_mon~y, ylab="Contacts Count 12 mon",xlab="Attrited Customer? (0=No, 1=Yes)")#s
boxplot(Credit_Limit~y, ylab="Credit_Limit",xlab="Attrited Customer? (0=No, 1=Yes)") #
boxplot(Total_Revolving_Bal~y, ylab="Total_Revolving_Bal",xlab="Attrited Customer? (0=No, 1=Yes)")#s
boxplot(Avg_Open_To_Buy~y, ylab="Avg Open To Buy",xlab="Attrited Customer? (0=No, 1=Yes)") #
boxplot(Total_Amt_Chng_Q4_Q1~y, ylab="Total Amt Chng Q4 Q1",xlab="Attrited Customer? (0=No, 1=Yes)")
boxplot(Total_Trans_Amt~y, ylab="Total Trans Amt",xlab="Attrited Customer? (0=No, 1=Yes)") #
boxplot(Total_Trans_Ct~y, ylab="Total_Trans_Ct",xlab="Attrited Customer? (0=No, 1=Yes)") #
boxplot(Total_Ct_Chng_Q4_Q1~y, ylab="Total Ct Chng Q4 Q1",xlab="Attrited Customer? (0=No, 1=Yes)")
boxplot(Avg_Utilization_Ratio~y, ylab="Avg Utilization Ratio",xlab="Attrited Customer? (0=No, 1=Yes)")#s
```

The distributions of some predictors are quite skewed. In this cases we include both predictor and log(predictor)/sqrt(predictor) as potential predictors in the logistic regression model

## Fit logistic Regression full model

```{r}
m2 <- glm(y~Gender+Dependent_count+Education_Level+Marital_Status+Income_Category+Card_Category+Customer_Age+Dependent_count+log(Months_on_book)
          +Months_on_book+Total_Relationship_Count+log(Total_Relationship_Count)+Months_Inactive_12_mon+Contacts_Count_12_mon+Credit_Limit
          +log(Credit_Limit)+Total_Revolving_Bal+Avg_Open_To_Buy+log(Avg_Open_To_Buy)+Total_Amt_Chng_Q4_Q1+Total_Trans_Amt+log(Total_Trans_Amt)
          +Total_Trans_Ct+log(Total_Trans_Ct)+Total_Ct_Chng_Q4_Q1+Avg_Utilization_Ratio,family=binomial(),data=bank_complete)
m_2 <- glm(y~Customer_Age+Dependent_count+log(Months_on_book)+Months_on_book+Total_Relationship_Count+log(Total_Relationship_Count)
          +Months_Inactive_12_mon+Contacts_Count_12_mon+Credit_Limit+log(Credit_Limit)+Total_Revolving_Bal+Avg_Open_To_Buy+log(Avg_Open_To_Buy)
          +Total_Amt_Chng_Q4_Q1+Total_Trans_Amt+log(Total_Trans_Amt)+Total_Trans_Ct+log(Total_Trans_Ct)+Total_Ct_Chng_Q4_Q1+Avg_Utilization_Ratio,family=binomial(),data=bank_complete)
```

### Obtaining marginal model plots

```{r}
mmps(m_2,layout=c(2,3))
```

```{r}
summary(m2)
```

## Adding Sqrt transformations

```{r}
m3 <- glm(y~Gender+Education_Level+Marital_Status+Income_Category+Card_Category+Customer_Age+Dependent_count+sqrt(Dependent_count)+log(Months_on_book)+
            Months_on_book+Total_Relationship_Count+log(Total_Relationship_Count)+Months_Inactive_12_mon+sqrt(Months_Inactive_12_mon)+
            Contacts_Count_12_mon+sqrt(Contacts_Count_12_mon)+Credit_Limit+log(Credit_Limit)+Total_Revolving_Bal+sqrt(Total_Revolving_Bal)+
            Avg_Open_To_Buy+log(Avg_Open_To_Buy)+Total_Amt_Chng_Q4_Q1+Total_Trans_Amt+log(Total_Trans_Amt)+Total_Trans_Ct+log(Total_Trans_Ct)
           +Total_Ct_Chng_Q4_Q1+Avg_Utilization_Ratio+sqrt(Avg_Utilization_Ratio),family=binomial(),data=bank_complete)
m_3 <- glm(y~Customer_Age+Dependent_count+sqrt(Dependent_count)+log(Months_on_book)+
            Months_on_book+Total_Relationship_Count+log(Total_Relationship_Count)+Months_Inactive_12_mon+sqrt(Months_Inactive_12_mon)+
            Contacts_Count_12_mon+sqrt(Contacts_Count_12_mon)+Credit_Limit+log(Credit_Limit)+Total_Revolving_Bal+sqrt(Total_Revolving_Bal)+
            Avg_Open_To_Buy+log(Avg_Open_To_Buy)+Total_Amt_Chng_Q4_Q1+Total_Trans_Amt+log(Total_Trans_Amt)+Total_Trans_Ct+log(Total_Trans_Ct)
           +Total_Ct_Chng_Q4_Q1+Avg_Utilization_Ratio+sqrt(Avg_Utilization_Ratio),family=binomial(),data=bank_complete)
```

### Output from R for m3

```{r}
summary(m3)
```


### Obtaining marginal model plots

```{r}
mmps(m_3,layout=c(2,3))
```

## Using the difference in deviance to compare the nested models m_2 and m3

```{r}
anova(m2,m3,test="Chisq")
```

The addition of the terms have significantly reduced the deviance so m3 is the best model

## Examine leverage values and standardized deviance residuals for m3

```{r}
par(mfrow=c(1,1))
hvalues <- influence(m3)$hat
stanresDeviance <- residuals(m3)/sqrt(1-hvalues)
plot(hvalues,stanresDeviance,ylab="Standardized Deviance Residuals",xlab="Leverage Values",ylim=c(-4,4),xlim=c(-0.03,0.7))
abline(v=2*26/length(y),lty=2)
```

## optimal model

### stepwise selection

```{r}
m3.step <- step(m3,direction="both")
summary(m3.step)
```

### Backward selection

```{r}
m3.back <- step(m3)
summary(m3.back)
```

### forward selection

```{r}
m3.null <- glm(y ~ 1 , family=binomial(link=logit) ,data=bank_complete)
m3.forward <- step(m3.null,
                         scope=list(lower=formula(m3.null),upper=formula(m3)), 
                         direction="forward",  data=bank_complete)
summary(m3.forward)
```

### Using the difference in deviance to compare among the three models

```{r}
anova(m3.step,m3.back,test="Chisq")
anova(m3.forward,m3.step,test="Chisq")
```

## prediction and ROC

```{r}
pred <- predict(m3.step, type="response")

par(mfrow=c(1,1))
m3.step.roc <- roc(y ~ pred, plot = TRUE, print.auc = TRUE)
```
```{r}
```

## Misclassification Rate:

```{r}
class.table <- data.frame(response = y, predicted = round(pred,0))
xtabs(~ predicted + response, data= class.table)
```
```{r}
miscl_Error = (253+517)/10127
miscl_Error
```

## Training and Test data - Dividing data randomly between Training and Test for prediction #avg open to buy and total revolving balance


```{r}
spl.split = sample.split(bank_complete$Attrition_Flag, SplitRatio = 0.02)
```

```{r}
train.spl = subset(bank_complete, spl.split == FALSE)
```

```{r}
test.spl = subset(bank_complete, spl.split == TRUE)
```

```{r}
m4 <- glm(train.spl$Attrition_Flag~Gender+Education_Level+Marital_Status+Income_Category+Card_Category+Customer_Age+Dependent_count+sqrt(Dependent_count)+log(Months_on_book)+
            Months_on_book+Total_Relationship_Count+log(Total_Relationship_Count)+Months_Inactive_12_mon+sqrt(Months_Inactive_12_mon)+
            Contacts_Count_12_mon+sqrt(Contacts_Count_12_mon)+Credit_Limit+log(Credit_Limit)+
            Avg_Open_To_Buy+log(Avg_Open_To_Buy)+Total_Amt_Chng_Q4_Q1+Total_Trans_Amt+log(Total_Trans_Amt)+Total_Trans_Ct+log(Total_Trans_Ct)
           +Total_Ct_Chng_Q4_Q1+Avg_Utilization_Ratio+sqrt(Avg_Utilization_Ratio),family=binomial(),data=train.spl)
```

```{r}
summary(m4)
```

### stepwise selection

```{r}
m4.step <- step(m4,direction="both")
summary(m4.step)
```

### Backward selection

```{r}
m4.back <- step(m4)
summary(m4.back)
```

### forward selection

```{r}
m4.null <- glm(train.spl$Attrition_Flag ~ 1 , family=binomial(link=logit) ,data=train.spl)
m4.forward <- step(m4.null,
                         scope=list(lower=formula(m4.null),upper=formula(m4)), 
                         direction="forward",  data=train.spl)
summary(m4.forward)
```
```{r}
anova(m4.step,m4.back,test="Chisq")
```
```{r}
anova(m4.forward,m4.step,test="Chisq")
```
```{r}
summary(m4.step)
                        
```
## Prediction and misclassification error for training data

```{r}
pred.train.spl = predict(m4.step, newdata = train.spl,
                      type = "response")
 
```

```{r}
class.table3 <- data.frame(response3 = train.spl$Attrition_Flag, predicted3 = round(pred.train.spl,0))
xtabs(~ predicted3 + response3, data= class.table3)
```

```{r}
miscl_Error_train = (249+505)/9924
miscl_Error_train
```

## Prediction and misclassification error for test data

```{r}
pred.test.spl = predict(m4.step, newdata = test.spl,
                      type = "response")
 
```

```{r}
class.table2 <- data.frame(response2 = test.spl$Attrition_Flag, predicted2 = round(pred.test.spl,0))
xtabs(~ predicted2 + response2, data= class.table2)
```

```{r}
miscl_Error_test = (12+3)/203
miscl_Error_test
```
